Disease Knowledge Transfer across Neurodegenerative Diseases

### 2019-10-04 paper/pres/poster-dkt.tex
more edits on dkt poster

\author{R\u{a}zvan V. Marinescu\inst{1,2}, Marco Lorenzi\inst{3}, Stefano B. Blumberg\inst{1}, Alexandra L. Young\inst{1}, Pere Planell-Morell\inst{1}, Neil P. Oxtoby\inst{1},\\ Arman Eshaghi\inst{1,4}, Keir X. Young\inst{5}, Sebastian J. Crutch\inst{5}, Polina Golland\inst{2}, Daniel C. Alexander\inst{1}\\
\begin{tabular}{l p{1cm} l p{1cm} l}
\large{$^1$Centre for Medical Image Computing, UCL, UK}  & & \large{$^3$University of C\^{o}te d'Azur, Inria Sophia Antipolis, France} & & \large{$^5$Dementia Research Centre, UCL, UK}\\
\large{$^2$Computer Science and Artificial Intelligence Laboratory, MIT, USA} & & \large{$^4$Queen Square MS Centre, UCL Institute of Neurology, UK} & & 
 \\
\end{tabular}
}

@@

\begin{minipage}{50cm}					%Minipage for title contents
% \vspace{-18cm}
\maketitle
\end{minipage}
\begin{textblock}{0}(10,0.2)
\begin{tabular}{l}
\includegraphics[height=5cm]{ucl_logo}\\
\includegraphics[height=5cm]{mit_logo}\\
\end{tabular}
\includegraphics[height=6cm]{qr_dkt.png}

@@

\newcommand{\heading}[1]{\textbf{\Huge{#1}}}

@@

\heading{Aim}: Infer progression of non-MRI biomarkers in rare neurodegenerative diseases by leveraging larger datasets of common neurodegenerative diseases.

\section*{Why}

\begin{itemize}
 \item Rare neurodegenerative diseases not well understood
 \item Identify outcome measures and subjects for clinical trials
\end{itemize}

@@

\section*{Challenges}

@@

\section*{Method}

\vspace{1em}
\begin{figure}[H]
 \centering
   \begin{subfigure}{0.49\columnwidth}
   \centering
   \fnt{1. Each disease characterised by region-specific dysfunction profile}\\
   $ \gamma_{ij}^l = f(\beta_{i} + m_{ij}; \lambda_{d_i}^l) $\\
%    \vspace{0.9em}
%    \includegraphics[width=\textwidth]{diseaseEvol}

@@

\heading{Inference}: Perform loopy belief propagation

@@

\begin{itemize}
 \item Dementia Research Center cohort: 76 PCA, 67 tAD, 87 age-matched controls.
 \item TADPOLE Challenge dataset split into three cohorts with different progressions: hippocampal, cortical and subcortical.
 \item Synthetic dataset mimicking the DRC cohort.
\end{itemize}

@@

\section*{Results}

\begin{figure}[H]
\fnt{In synthetic experiment, the estimated parameters are close to the true parameters}\\
\includegraphics[width=\columnwidth]{../figures/compTrueParams105_synth1_JMD.pdf}
  \label{fig:dktSynthTrajCompTrue}
\end{figure}

\begin{figure}[H]
% \centering
\fnt{Inferred multimodal trajectories for Posterior Cortical Atrophy are plausible, suggesting late-stage parietal and occipital damage. }\\
 \includegraphics[width=\columnwidth, trim=0 20 0 0, clip]{../figures/trajDisSpaceOverlap_PCA_tad-drcTinyPen5_JMD.png}
%  \caption{Estimated trajectories for the PCA cohort. The only data that were available were the MRI volumetric data. The dynamics of the other biomarkers has been inferred by the model using data from typical AD, and taking into account the different spatial distribution of pathology in PCA vs tAD. }
%  \label{fig:PCAtrajByModality}
\end{figure}

\fnt{Our model has favourable performance compared to other models, on two different datasets}\\
\begin{figure}[H]
\fontsize{27}{30}\selectfont
\begin{tabular}{c | c c c c c c}
\textbf{Model} & \textbf{Cingulate} & \textbf{Frontal} & \textbf{Hippocam.} & \textbf{Occipital} & \textbf{Parietal} & \textbf{Temporal}\\
& \multicolumn{6}{c}{\textbf{TADPOLE: Hippocampal subgroup to Cortical subgroup}}\\
DKT (ours) &      0.56 $\pm$ 0.23 &    \textbf{0.35 $\pm$ 0.17} &        \textbf{0.58 $\pm$ 0.14} &     -0.10 $\pm$ 0.29 &     \textbf{0.71 $\pm$ 0.11} &     \textbf{0.34 $\pm$ 0.26} \\
Latent stage &      0.44 $\pm$ 0.25 &    0.34 $\pm$ 0.21 &       0.34 $\pm$ 0.24* &     \textbf{-0.07 $\pm$ 0.22} &     0.64 $\pm$ 0.16 &    0.08 $\pm$ 0.24* \\
Multivariate &      \textbf{0.60 $\pm$ 0.18} &   0.11 $\pm$ 0.22* &       0.12 $\pm$ 0.29* &     -0.22 $\pm$ 0.22 &   -0.44 $\pm$ 0.14* &   -0.32 $\pm$ 0.29* \\
Spline &    -0.24 $\pm$ 0.25* &  -0.06 $\pm$ 0.27* &        0.58 $\pm$ 0.17 &     -0.16 $\pm$ 0.27 &    0.23 $\pm$ 0.25* &    0.10 $\pm$ 0.25* \\
Linear &    -0.24 $\pm$ 0.25* &   0.20 $\pm$ 0.25* &        0.58 $\pm$ 0.17 &     -0.16 $\pm$ 0.27 &    0.23 $\pm$ 0.25* &    0.13 $\pm$ 0.23* \\
& \multicolumn{6}{c}{\textbf{typical Alzheimer's to Posterior Cortical Atrophy}}\\
DKT (ours) &    0.77 $\pm$ 0.11 &    0.39 $\pm$ 0.26 &      0.75 $\pm$ 0.09 &    0.60 $\pm$ 0.14 &    \textbf{0.55 $\pm$ 0.24} &    \textbf{0.35 $\pm$ 0.22} \\
Latent stage &    \textbf{0.80 $\pm$ 0.09} &    \textbf{0.53 $\pm$ 0.17} &      \textbf{0.80 $\pm$ 0.12} &    0.56 $\pm$ 0.18 &    0.50 $\pm$ 0.21 &    0.32 $\pm$ 0.24 \\
Multivariate &   0.73 $\pm$ 0.09 &   0.45 $\pm$ 0.22  &    0.71 $\pm$ 0.08 & -0.28 $\pm$ 0.21* &  0.53 $\pm$ 0.22  &  0.25 $\pm$ 0.23* \\
Spline &   0.52 $\pm$ 0.20* &  -0.03 $\pm$ 0.35* &     0.66 $\pm$ 0.11* &   0.09 $\pm$ 0.25* &    0.53 $\pm$ 0.20 &   0.30 $\pm$ 0.21* \\
Linear &   0.52 $\pm$ 0.20* &    0.34 $\pm$ 0.27 &     0.66 $\pm$ 0.11* &    \textbf{0.64 $\pm$ 0.17} &    0.54 $\pm$ 0.22 &   0.30 $\pm$ 0.21* \\
\end{tabular}
\vspace{0.5em}
% \caption[Performance evaluation of DKT and other models]{Performance evaluation of DKT and four other statistical models of decreasing complexity. }
% \end{footnotesize}
\label{sec:dktPerfMetrics}
\end{figure}

\section*{Conclusion}

\begin{itemize}
 \item Developed a novel methodology and model for transfer learning across different diseases
%  \item Inferred, likely for the first time, continuous multimodal trajectories for Posterior Cortical Atrophy
%  \item 
\end{itemize}

@@

\subsection*{Weblinks}
\begin{itemize}
\item Source code: \url{https://github.com/mrazvan22/dkt}
\item Personal website: \url{https://people.csail.mit.edu/razvan/}
\end{itemize}
}

% \vspace{}

@@

\subsection*{Funders and Grants}

% \begin{figure}[H]
% \begin{subfigure}{0.48\columnwidth}
\vspace{1em}
% \hspace{2em}
\includegraphics[height=4.0cm]{epsrc_logo}
\hspace{0.5em}
\includegraphics[height=4cm]{nac_logo} 
\hspace{0.5em}
\includegraphics[height=4cm]{cdt_logo} 
\hspace{0.5em}
\includegraphics[height=4cm]{nih_logo} 
\hspace{0.5em}
\includegraphics[height=4cm]{aruk_logo} 
\hspace{0.5em}
% \end{subfigure}
% \begin{subfigure}{0.48\columnwidth}
%  \begin{itemize}
%  \item  EPSRC grant EP/L016478/1
%  \item  NIH grant NIBIB NAC P41EB015902
% \end{itemize}
% \end{subfigure}

>>>>

\author{\LARGE{R\u{a}zvan Marinescu\inst{1,2}, Marco Lorenzi\inst{3}, Stefano Blumberg\inst{1}, Alexandra Young\inst{1}, Pere Morell\inst{1}},\\ \LARGE{Neil Oxtoby\inst{1}, Arman Eshaghi\inst{1,4}, Keir Young\inst{5}, Sebastian Crutch\inst{5}, Polina Golland\inst{2}, Daniel Alexander\inst{1}}\\
\begin{tabular}{l p{1cm} l p{1cm} l}
\large{$^1$Centre for Medical Image Computing, UCL, UK}  & & \large{$^3$University of C\^{o}te d'Azur, Inria Sophia Antipolis, France} & & \large{$^5$Dementia Research Centre, UCL, UK}\\
\large{$^2$Computer Science and Artificial Intelligence Laboratory, MIT, USA} & & \large{$^4$Queen Square MS Centre, UCL Institute of Neurology, UK} & & 
 \\
\end{tabular}
}

@@

\begin{minipage}{50cm}					%Minipage for title contents
% \vspace{-18cm}
\maketitle
\end{minipage}
\begin{textblock}{0}(12.2,0.2)
\begin{tabular}{l}
\includegraphics[height=5.5cm]{mit_logo}\includegraphics[height=6cm]{qr_dkt.png}\\
\includegraphics[height=5cm]{ucl_logo}\\
\end{tabular}

@@

\newcommand{\heading}[1]{\noindent\textbf{\Huge{#1}}}

@@

\heading{Aim} Infer progression of non-MRI biomarkers in rare neurodegenerative diseases (NDs) by leveraging larger datasets of common NDs.

\heading{Why} Posterior Cortical Atrophy (PCA): progression of non-MRI biomarkers not known $\rightarrow$ Identify outcome measures and suitable subjects for PCA clinical trials

@@

\heading{Challenges}

@@

\heading{Method}

\vspace{0.6em}
\begin{figure}[H]
 \centering
   \begin{subfigure}{0.49\columnwidth}
   \centering
   \fnt{1. Each disease characterised by region-specific dysfunction profile}\\
   $ \gamma_{ij}^l = f(\beta_{i} + m_{ij}; \lambda_{d_i}^l) $\\
%    \vspace{0.9em}
%    \includegraphics[width=\textwidth]{diseaseEvol}

@@

\heading{Inference:} Loopy belief propagation

@@

\begin{itemize}
 \item Dementia Research Center cohort: 76 PCA, 67 tAD, 87 age-matched controls for training, 10 PCA with DTI for validation
 \item TADPOLE Challenge dataset split into three subgroups with different progressions: 21 hippocampal, 35 cortical, 27 subcortical.
 \item Synthetic dataset mimicking the DRC cohort: 50 subjects with "synthetic PCA", 100 subjects with "synthetic AD".
\end{itemize}

@@

\heading{Results}

\vspace{0.6em}
\begin{figure}[H]
\fnt{In synthetic experiment, the estimated parameters are close to the true parameters}\\
\includegraphics[width=\columnwidth]{../figures/compTrueParams105_synth1_JMD.pdf}
  \label{fig:dktSynthTrajCompTrue}
\end{figure}

\begin{figure}[H]
% \centering
\fnt{Inferred multimodal trajectories for Posterior Cortical Atrophy are plausible, suggesting late-stage parietal and occipital damage. }
 \includegraphics[width=\columnwidth, trim=0 20 0 35, clip]{../figures/trajDisSpaceOverlap_PCA_tad-drcTinyPen5_JMD.png}
%  \caption{Estimated trajectories for the PCA cohort. The only data that were available were the MRI volumetric data. The dynamics of the other biomarkers has been inferred by the model using data from typical AD, and taking into account the different spatial distribution of pathology in PCA vs tAD. }
%  \label{fig:PCAtrajByModality}
\end{figure}

\noindent \fnt{Our model has favourable performance compared to other models, on two different datasets}
\vspace{0.3em}
\begin{figure}[H]
\fontsize{27}{30}\selectfont
\begin{tabular}{c | c c c c c c}
\textbf{Model} & \textbf{Cingulate} & \textbf{Frontal} & \textbf{Hippocam.} & \textbf{Occipital} & \textbf{Parietal} & \textbf{Temporal}\\
& \multicolumn{6}{c}{\textbf{TADPOLE: Hippocampal subgroup to Cortical subgroup}}\\
DKT (ours) &      0.56 $\pm$ 0.23 &    \textbf{0.35 $\pm$ 0.17} &        \textbf{0.58 $\pm$ 0.14} &     -0.10 $\pm$ 0.29 &     \textbf{0.71 $\pm$ 0.11} &     \textbf{0.34 $\pm$ 0.26} \\
Latent stage &      0.44 $\pm$ 0.25 &    0.34 $\pm$ 0.21 &       0.34 $\pm$ 0.24* &     \textbf{-0.07 $\pm$ 0.22} &     0.64 $\pm$ 0.16 &    0.08 $\pm$ 0.24* \\
Multivariate &      \textbf{0.60 $\pm$ 0.18} &   0.11 $\pm$ 0.22* &       0.12 $\pm$ 0.29* &     -0.22 $\pm$ 0.22 &   -0.44 $\pm$ 0.14* &   -0.32 $\pm$ 0.29* \\
Spline &    -0.24 $\pm$ 0.25* &  -0.06 $\pm$ 0.27* &        0.58 $\pm$ 0.17 &     -0.16 $\pm$ 0.27 &    0.23 $\pm$ 0.25* &    0.10 $\pm$ 0.25* \\
Linear &    -0.24 $\pm$ 0.25* &   0.20 $\pm$ 0.25* &        0.58 $\pm$ 0.17 &     -0.16 $\pm$ 0.27 &    0.23 $\pm$ 0.25* &    0.13 $\pm$ 0.23* \\
& \multicolumn{6}{c}{\textbf{typical Alzheimer's to Posterior Cortical Atrophy}}\\
DKT (ours) &    0.77 $\pm$ 0.11 &    0.39 $\pm$ 0.26 &      0.75 $\pm$ 0.09 &    0.60 $\pm$ 0.14 &    \textbf{0.55 $\pm$ 0.24} &    \textbf{0.35 $\pm$ 0.22} \\
Latent stage &    \textbf{0.80 $\pm$ 0.09} &    \textbf{0.53 $\pm$ 0.17} &      \textbf{0.80 $\pm$ 0.12} &    0.56 $\pm$ 0.18 &    0.50 $\pm$ 0.21 &    0.32 $\pm$ 0.24 \\
Multivariate &   0.73 $\pm$ 0.09 &   0.45 $\pm$ 0.22  &    0.71 $\pm$ 0.08 & -0.28 $\pm$ 0.21* &  0.53 $\pm$ 0.22  &  0.25 $\pm$ 0.23* \\
Spline &   0.52 $\pm$ 0.20* &  -0.03 $\pm$ 0.35* &     0.66 $\pm$ 0.11* &   0.09 $\pm$ 0.25* &    0.53 $\pm$ 0.20 &   0.30 $\pm$ 0.21* \\
Linear &   0.52 $\pm$ 0.20* &    0.34 $\pm$ 0.27 &     0.66 $\pm$ 0.11* &    \textbf{0.64 $\pm$ 0.17} &    0.54 $\pm$ 0.22 &   0.30 $\pm$ 0.21* \\
\end{tabular}
\vspace{0.5em}
% \caption[Performance evaluation of DKT and other models]{Performance evaluation of DKT and four other statistical models of decreasing complexity. }
% \end{footnotesize}
\label{sec:dktPerfMetrics}
\end{figure}

\heading{Conclusion}

\begin{itemize}
 \item Developed a novel methodology and model for transfer learning across different diseases
 \item Inferred multimodal trajectories for Posterior Cortical Atrophy
%  \item 
\end{itemize}

@@

\vspace{-1em}
\subsection*{Weblinks}
\begin{itemize}
\item Source code: \url{https://github.com/mrazvan22/dkt}
\item Website: \url{https://people.csail.mit.edu/razvan/}
\end{itemize}
}

@@

\subsection*{Funders}

% \begin{figure}[H]
% \begin{subfigure}{0.48\columnwidth}
\vspace{0em}
% \hspace{2em}
\newcommand{\heightLogos}{3.5cm}
\includegraphics[height=\heightLogos]{epsrc_logo}
\hspace{0.5em}
\includegraphics[height=\heightLogos]{nac_logo} 
\hspace{0.5em}
\includegraphics[height=\heightLogos]{cdt_logo} 
\hspace{0.5em}
\includegraphics[height=\heightLogos]{nih_logo} 
\hspace{0.5em}
\includegraphics[height=\heightLogos]{aruk_logo} 
\hspace{0.5em}
% \end{subfigure}
% \begin{subfigure}{0.48\columnwidth}
%  \begin{itemize}
%  \item  EPSRC grant EP/L016478/1
%  \item  NIH grant NIBIB NAC P41EB015902
% \end{itemize}
% \end{subfigure}
